---
title: "Lab 6"
author: "Frank Woodling"
date: "March 29, 2016"
output: pdf_document
---

1. 

```{r}
list <- c(92, 123, 114, 72, 117, 111, 129, 113, 102, 98, 94, 96)
names(list) <- c("PG-13", "PG-13", "PG-13", "G", "PG-13", "R", "R", "PG","R","PG","PG", "R")

list.sorted <- sort(list)
rbind(list.sorted, 1:12)

g.ranks <- c(1)
pg.ranks <- c(3, 5, 8)
pg13.ranks <- c(2, 9, 10, 11)
r.ranks <- c(4, 6, 7, 12)

# group         ranks         sample size         mean rank
# 
# G              12                1                 12
# 
# PG           3, 5, 8             3                 5.33
# 
# PG-13      2, 9, 10, 11          4                 8  
# 
# R          4, 6, 7, 12           4                 7.25   

kw.1 <- 1*(12-13/2)^2
kw.2 <- 3*(5.33-13/2)^2
kw.3 <- 4*(8-13/2)^2
kw.4 <- 4*(7.25-13/2)^2

kw <- (12/(12*13))*(kw.1+kw.2+kw.3+kw.4)
kw
```

The Kruskal-Wallis statistic is 3.50641.


# 2.
```{r}
expsum.g <- sum(1:12)*1/12
expsum.pg <- sum(1:12)*3/12
expsum.pg13 <- sum(1:12)*4/12
expsum.r <- sum(1:12)*4/12

# group         ranks         observed rank-sum        expected rank-sum
# 
# G              12                12                         6.5
# 
# PG           3, 5, 8             16                        19.5
# 
# PG-13      2, 9, 10, 11          32                        26 
# 
# R          4, 6, 7, 12           29                        26 

kw2.1 <- ((12-6.5)^2)/6.5
kw2.2 <- ((16-19.5)^2)/19.5
kw2.3 <- ((32-26)^2)/26
kw2.4 <- ((29-26)^2)/26

kw2 <- 6/12*(kw2.1 + kw2.2 + kw2.3 + kw2.4)
kw2
```

The Kruskal-Wallis statistic is 3.50641 in this test as well.

# 3. 
```{r}
list.2 <- c(95, 97, 100, 95, 96, 101, 102, 100, 141, 87, 154, 107)
names(list.2) <- c("PG-13", "R", "G", "PG", "PG-13", "PG-13", "R", "PG-13", "PG-13", "PG", "R", "PG-13")

list.sorted2 <- sort(list.2)
rbind(list.sorted2, 1:12)

# group     ranks                       sample size         mean rank
# 
# G         6.5                             1                  6.5 
# 
# PG        1, 2.5                          2                  1.75
# 
# PG-13     2.5, 4, 6.5, 8, 10, 11          6                  7
# 
# R         5, 9, 12                        3                  8.666667

kw3.1 <- 1*(6.5-13/2)^2
kw3.2 <- 2*(1.75-13/2)^2
kw3.3 <- 6*(7-13/2)^2
kw3.4 <- 3*(8.66667-13/2)^2
kw3 <- (12/(12*13))*(kw3.1+kw3.2+kw3.3+kw3.4)
kw3

# there are 2 ties, 95 and 100 with two a piece
kwties.denom <- 1-(2*(2^3-2))/(12^2-12)
kw3.ties <- kw3/kwties.denom
kw3.ties
```


# 4.
```{r}
movies <- read.table("moviesall.txt", header=T)
attach(movies)
table(runtime)

### calculate the observed KW statistic
t.j = c(2,6,3,3,2,3,5,2,3,4,5,3,3,5,4,5,4,3,6,2,3,4,3,3,3,4,5,3,2,3,2,2,2,3)
n.i = c(4, 21, 65, 50); N = sum(n.i)
ranks = rank(runtime) ### rank the data
R.i = c(mean(ranks[rating=="G"]), mean(ranks[rating=="PG"]), mean(ranks[rating=="PG-13"]), 
        mean(ranks[rating=="R"]))

KW.noties = 12/(N*(N+1)) * sum( n.i*(R.i - (N+1)/2)^2 )
teststat.obs = KW.noties/( 1 - sum( t.j^3 - t.j )/(N^3 - N) )
teststat = rep(NA, 1000)

for(i in 1:1000) {

### randomly "shuffle" the rating labels for the movies
ratingSHUFFLE = sample(rating)

### compute the KW statistic for the shuffled data
R.i = c(mean(ranks[ratingSHUFFLE=="G"]), mean(ranks[ratingSHUFFLE=="PG"]), 
        mean(ranks[ratingSHUFFLE=="PG-13"]), mean(ranks[ratingSHUFFLE=="R"]))
KW.noties = 12/(N*(N+1)) * sum( n.i*(R.i - (N+1)/2)^2 )
teststat[i] = KW.noties/( 1 - sum( t.j^3 - t.j )/(N^3 - N) )
}

### calculate the approximate p-value
sum(teststat >= teststat.obs)/1000
teststat.obs

hist(table(runtime), main="Histogram of frequency of movie runtime")
```
Looking at the table and histogram we can see that 6 is the maximum number of tied observations on 84 minutes and 105 minutes.  It seems like the longer a movie is the less chance of a tie, and most of the ties are for lower runtimes.

# 5. 


# 6. 
```{r}
sum(teststat >= teststat.obs)/1000
teststat.obs
```
The p-value is 0 and the observed test statistic is 19.67098.  I seem to get a different number each time I try this.  I also got 0.001 and 6700 for the test statistic.

# 7.
```{r}
kruskal.test(runtime~rating, data = movies)
```
This value lines up with the lines of code in part 4.

# Apply what you have learned here
# 1.
```{r}
movies <- read.table("moviesall.txt", header=T)
attach(movies)

### calculate the observed KW statistic
t.j = c(2,6,3,3,2,3,5,2,3,4,5,3,3,5,4,5,4,3,6,2,3,4,3,3,3,4,5,3,2,3,2,2,2,3)
n.i = c(4, 21, 65, 50); N = sum(n.i)
ranks = rank(score) ### rank the data
R.i = c(mean(ranks[rating=="G"]), mean(ranks[rating=="PG"]), mean(ranks[rating=="PG-13"]), 
        mean(ranks[rating=="R"]))

KW.noties = 12/(N*(N+1)) * sum( n.i*(R.i - (N+1)/2)^2 )
teststat.obs = KW.noties/( 1 - sum( t.j^3 - t.j )/(N^3 - N) )
teststat = rep(NA, 1000)

for(i in 1:1000) {

### randomly "shuffle" the rating labels for the movies
ratingSHUFFLE = sample(rating)

### compute the KW statistic for the shuffled data
R.i = c(mean(ranks[ratingSHUFFLE=="G"]), mean(ranks[ratingSHUFFLE=="PG"]), 
        mean(ranks[ratingSHUFFLE=="PG-13"]), mean(ranks[ratingSHUFFLE=="R"]))
KW.noties = 12/(N*(N+1)) * sum( n.i*(R.i - (N+1)/2)^2 )
teststat[i] = KW.noties/( 1 - sum( t.j^3 - t.j )/(N^3 - N) )
}

### calculate the approximate p-value
sum(teststat >= teststat.obs)/1000
teststat.obs

kruskal.test(score~rating)
```
We fail to reject the null hypothesis with a p-value of 0.541 (the built in test says 0.5279).  We can conclude that there is a difference in means between the scores given for a movie and the rating (G, PG, PG-13, R).  

# 2.
```{r}
kruskal.test(gross~rating)
```
The null hypothesis is that there is not a difference of means between the different ratings and their box office gross.  In this case we fail to reject the null hypothesis with a p-value of 0.07781.  We can conclude that the is not a significant difference between the mean box office gross throughout the different ratings. 


# Test
```{r}
# 1
tlist <- c(92, 123, 114, 72, 117, 111, 129, 113, 102, 98, 94, 96)
tnames <- c("PG-13", "PG-13", "PG-13", "G", "PG-13", "R", "R", "PG","R","PG","PG", "R")
tgroups <- c(3,3,3,1,3,4,4,2,4,2,2,4)
kruskal.test(tlist, tgroups)


# 3
tlist.3 <- c(95, 97, 100, 95, 96, 101, 102, 100, 141, 87, 154, 107)
tgroups.3 <- c(3, 4, 1, 2, 3, 3, 4, 3, 3, 2, 4, 3)
kruskal.test(tlist.3, tgroups.3)
```

