---
title: 'Lab 8: Chi-square'
author: "Frank Woodling"
date: "April 27, 2016"
output: pdf_document
---

# 1.
$H_{0}:$ Boys and girls prefer the same types of sports

$H_{1}:$ Boys and girls do not prefer the same types of sports

# 2.
No, from page 27 of the chapter 5 lecture slides the chi-square distribution may not be valid when some $e_{ij}<5$. We can still create a permutation distribution of the chi-square statistic though.

# 3.
The boys row currently adds up to $\frac{3}{5}$, and the girls row adds to $\frac{2}{5}$.  Both these would need to equal 1 in order for a traditional test to work properly.  We we need at least another 5 students total, 2 boys and 3 girls.  

# 4. 
```{r}
source("http://www4.stat.ncsu.edu/~lu/ST505/Rcode/functions-Ch5.R")

sport = c(1,2,2,1,3)
gender = c(1,1,1,2,2)

partytable <- table(gender, sport)
result_obs = chisq.test(partytable)
result_obs$statistic
```

# 5.
There are $\binom{5}{3}$ combinations or 10 total combinations.

```{r}
# n	    boys       girls
# 1	 B1, B2, F1	   F2, O1
# 2	 B1, B2, F2	   F1, O1
# 3	 B1, B2, O1	   F1, F2
# 4	 B1, F1, F2	   B2, O2
# 5	 B1, F1, O1	   B2, F2
# 6	 B1, F2, O1	   B2, F1
# 7	 B1, F1, F2	   B2, O1
# 8	 B2, F1, O1	   B1, F2
# 9	 B2, F2, O1	   B1, F1
# 10 F1, F2, O1    B1, B2
```


# 6.
```{r}
a <- matrix(
  c(2, 1,	0, 0,	1, 1,
    2, 1, 0, 0 ,1, 1,
    2, 0 ,1, 0, 2, 0,
    1, 2, 0, 1, 0, 1,
    1, 1, 1, 1, 1, 0,
    1, 1, 1, 1, 1, 0,
    1, 2, 0, 1, 0, 1,
    1, 1, 1, 1, 1, 0,
    1, 1, 1, 1, 1, 0,
    0, 2, 1, 2, 0, 0),
  nrow = 10,
  ncol = 6, 
  byrow = T)

dimnames(a) = list(1:10, c("B", "F", "O", "B", "F", "O"))
a
```

# 7. 
```{r}
table_chi = rep(0, 10)
for(i in 1:10)
{
  r1 = a[i, 1:3]
  r2 = a[i, 4:6]
  table = rbind(r1, r2)
  chi_result = chisq.test(table)
  table_chi[i] = chi_result$statistic
  
}

cbind(a, table_chi)
```

# 8. 
```{r}
mean(table_chi >= result_obs$statistic)
```
This seems like a very high p-value, and may be incorrect.  If 0.6 is the p-value then we fail to reject the null hypothesis.  We can conclude that boys and girls like the same types of sports.

# 9.
```{r}
### Make observed contingency table and calculate stat
Row1 = c(22,40,12); Row2 = c(30,17,11)
Table = rbind(Row1,Row2)
teststat.obs = chisq.test(Table)$statistic
teststat.obs
### create the prefernce data and the gender data
preference = c( rep("B",52), rep("F",57), rep("O",23))
gender = c( rep("boy",74), rep("girl",58) )
table(preference); table(gender)
y = preference; x = gender
teststat = rep(NA, 1000)
for(i in 1:1000) {
### randomly "shuffle" the y data between the x groups
ySHUFFLE = sample(y)
### compute chi-square stat for the shuffled data
TableSHUFFLE = table(x,ySHUFFLE)
teststat[i] = chisq.test(TableSHUFFLE)$statistic
}
### calculate the approximate p-value
mean(teststat >= teststat.obs)
```
I found a p-value of 0.08 with the full dataset. We can reject the null hypothesis, and conclude that boys and girls prefer different sports.


# 10.
```{r}
Row1 = c(22,40,12); Row2 = c(30,17,11)
Table = rbind(Row1,Row2)

chisq.test(Table)
```
Yes, since $e_{ij}>5$ we could use a traditional test.  The test resulted in a p-value of 0.01263 so we can reject the null hypothesis.  This value is much smaller than the permutation result.



# Summary
# 1.
The data set contains all of the data for boys and girls preferences of sports between basketball, football, or other sports.  For the permutation test we found a test-statistic of 8.74 and a p-value of 0.08, while in the traditional test we found a test-statistic of 8.744 and a p-value of 0.01263.  In both cases we can reject the null hypothesis, and conclude boys and girls prefer different sports.   

```{r, echo=F}
hist(teststat, main="Histogram of the Test Statistics")
```

# 2.
```{r}
#	     boys	            girls
# 1	  B1, B2, F1      	F2, O1
# 2	  B1, B2, F2	      F1, O1
# 3	  B1, B2, Soccer	  F1, F2
# 4	  B1, F1, F2	      B2, O2
# 5   B1, F1, Soccer   	B2, F2
# 6	  B1, F2, Soccer	  B2, F1
# 7	  B1, F1, F2	      B2, O1
# 8	  B2, F1, Soccer	  B1, F2
# 9	  B2, F2, Soccer	  B1, F1
# 10	F1, F2, Soccer	  B1, B2
# 11	B1, B2, F1	      F2, O1
# 12	B1, B2, F2	      F1, O1
# 13	B1, B2, Swimming	F1, F2
# 14	B1, F1, F2	      B2, O2
# 15	B1, F1, Swimming	B2, F2
# 16	B1, F2, Swimming	B2, F1
# 17	B1, F1, F2	      B2, O1
# 18	B2, F1, Swimming	B1, F2
# 19	B2, F2, Swimming	B1, F1
# 20	F1, F2, Swimming	B1, B2
# 21	B1, B2, F1	      F2, O1
# 22	B1, B2, F2	      F1, O1
# 23	B1, B2, Track   	F1, F2
# 24	B1, F1, F2	      B2, O2
# 25	B1, F1, Track    	B2, F2
# 26	B1, F2, Track	    B2, F1
# 27	B1, F1, F2	      B2, O1
# 28	B2, F1, Track    	B1, F2
# 29	B2, F2, Track	    B1, F1
# 30	F1, F2, Track	    B1, B2

```

I believe this would create a total of 30 combinations since there is only 1 other in each permutation between the boys and girls.  Each other could be split into the 3 new sports.