---
title: 'Lab 10: Bootstrap Bonus Lab'
author: "Frank Woodling"
date: "May 12, 2016"
output: pdf_document
---

# 1.
```{r}
GPA <- read.delim("C:/Users/Frank/Desktop/STAT 3480/Lab 10/GPA.txt")


attach(GPA)

summary(lm(CollGPA~SAT))
```
The equattion is Y = 0.0018*SAT + 0.1520. For each 0.00018020 increase in the SAT score their GPA will increase by 1.

# 2.
```{r}
### create our data and calculate thetahat, the slope of the regression line
oursample = GPA
thetahat = lm(CollGPA ~ SAT, data=oursample)$coeff[2]
thetahat
thetahat.b = rep(NA,1000)
for (i in 1:1000) {
### draw the bootstrap sample and calculate thetahat.b
index = 1:100
bootindex = sample(index, 100, replace=T)
bootsample = oursample[bootindex,]
thetahat.b[i] = lm(CollGPA ~ SAT, data=bootsample)$coeff[2]
}


### draw the bootstrap sample
index = 1:100
bootindex = sample(index, 100, replace=T)
bootsample = oursample[bootindex,]
bootsample
```
We can look at the bootindex and seee if the same index is repeated twice.  In my case I see both 1 and 6 repeated.

# 3.
```{r}
hist(thetahat.b)
quantile(thetahat.b, .025); quantile(thetahat.b, .975)
```
Out of an infinite amount of confidence intervals 95% of bootstrap intervals will fall between 0.001126614 and 0.002549624.

# 4.
It would be equal to .05, so 95% of intervals would bracket the true mean.

# 5.
```{r}
quantile(thetahat.b, .005); quantile(thetahat.b, .995)
```
The interval is between 0.0009297782 and 0.002757923 so the p-value would have to be less than 0.01.  We can say than of an infinite amount of bootstraps 99% are between that interval.

# 6.
It would have to contain 99% if all sample's means to be a 99% confidence interval.


# 7.
```{r}
detach(GPA)
GPAfull <- read.csv("C:/Users/Frank/Desktop/STAT 3480/Lab 10/GPA_full.txt", sep="")
attach(GPAfull)

summary(lm(CollGPA ~ SAT + HSGPA))
```
I am getting some strange regression results.  It lists 30 different coefficients for SAT.  No matter how I change it there are a ton of different SAT coefficients.


# 8. 
```{r}
### create our data
oursample = GPAfull
SATthetahat = lm(CollGPA ~ SAT + HSGPA + Rec, data=oursample)$coeff[2]
HSGPAthetahat = lm(CollGPA ~ SAT + HSGPA + Rec, data=oursample)$coeff[3]
Recthetahat = lm(CollGPA ~ SAT + HSGPA + Rec, data=oursample)$coeff[4]
SATthetahat; HSGPAthetahat; Recthetahat
SATthetahat.b = rep(NA,1000); HSGPAthetahat.b = rep(NA,1000); Recthetahat.b = rep(NA,1000)
for (i in 1:1000) {
### draw the bootstrap sample and calculate thetahat.b
index = 1:100
bootindex = sample(index, 100, replace=T)
bootsample = oursample[bootindex,]
SATthetahat.b[i] = lm(CollGPA ~ SAT + HSGPA + Rec, data=bootsample)$coeff[2]
HSGPAthetahat.b[i] = lm(CollGPA ~ SAT + HSGPA + Rec, data=bootsample)$coeff[3]
Recthetahat.b[i] = lm(CollGPA ~ SAT + HSGPA + Rec, data=bootsample)$coeff[4]
}
par(mfrow=c(1,3))
hist(SATthetahat.b); hist(HSGPAthetahat.b); hist(Recthetahat.b)
quantile(SATthetahat.b, .025); quantile(SATthetahat.b, .975)
quantile(HSGPAthetahat.b, .025); quantile(HSGPAthetahat.b, .975)
quantile(Recthetahat.b, .025); quantile(Recthetahat.b, .975)
```

# 9.
In only one interval does the confidence interval bracket 0.  

# 10.
```{r}
quantile(SATthetahat.b, .025);
quantile(SATthetahat.b, .975)
quantile(HSGPAthetahat.b, .025); quantile(HSGPAthetahat.b, .975)
quantile(Recthetahat.b, .025); 
quantile(Recthetahat.b, .975)
```

# This lab does not give enough information to answer this problem.  As we reduce the confidence level our confidence interval decreases.


# Summary
```{r}
fit = lm(CollGPA ~ SAT + HSGPA + Rec)
summary(fit)
predict(fit, GPAfull, interval="confidence") 
```
SAT scores, high school GPA, and number of positive recommendation letters are all positive indicators of college GPA.


CI(lm(CollGPA ~ SAT + HSGPA + Rec))



