---
title: "Homework 2"
author: "Frank Woodling"
date: "February 23, 2016"
output: pdf_document

---

# 2.01.
```{r}
library('gtools')

t1 <- c(1, 3, 5, 7)
t2 <- c(2, 4, 6, 8)

t1 <- c(2, 4, 6, 11)
t2 <- c(3, 5, 7, 12)
t = c(t1, t2)

teststat.obs = sum(t1)
teststat.obs
combo = combinations(8, 4, v = t, set = F, repeats.allowed = F)

W = NULL

for (i in 1:70)
{
  W = c(W, sum(combo[i,]))
}

sum(teststat.obs >= W)/choose(8, 4)
```
Before working out I thought I should get 0.33 or 0.5 for my p-value.  After plugging some numbers in I ended up with 0.3429 the first time and 0.3714 the second time.  

# 2.05 
```{r}
# library('gtools')
woodland <- read.table('woodland.txt', header=T)
woodland

attach(woodland)
ranks = rank(obs)

w.obs = sum(ranks[species=="A"])
a.ranks = combinations(10, 5, v = ranks, set = F, repeats.allowed = F)
W = rep(NA, choose(10,5))

for(i in 1:choose(10,5))
{
  W[i] = sum(a.ranks[i,])
}

sum(W > w.obs)/choose(10, 5)

```
With a p-value of 0.01587302 we can reject the null hypothesis that there is no difference between the two species' nesting heights.


# 2.12 complete first part, hodges-lehmann wrong (typo?)

```{r}
# From the table in the book the Hodges-Lehmann estimate should be from pairwise
# difference 5 to 25.

carapace <- read.table('carapace.txt', header=T)
carapace
attach(carapace)

s1 = c(5, 11, 16, 8, 12)
s2 = c(17, 14, 15, 21, 19, 13)

pwd = NULL

for(i in 1:5)
{
  for(j in 1:6)
  {
    pwd = c(pwd, (i-j))
  }
}

sort(pwd) 

hodges_ci = c(pwd[5], pwd[25])
hodges_ci

m = 5
n = 6
std_err = sqrt(((m-1)*sd(s1)+(n-1)*sd(s2))/(n+m-2))
t = abs(qt(.05, 9, lower.tail = TRUE))
upper = (mean(s1)-mean(s2)) + t*std_err*sqrt((1/m)+(1/n))
lower = (mean(s1)-mean(s2)) - t*std_err*sqrt((1/m)+(1/n))
upper
lower


```
The Hodge-Lehmann confidence interval is (-4, 4).  The normal confidence interval is (-8, -4).  I cannot figure out why these are not giving me reasonable confidence intervals.  I believe if anything it must be a problem with my loop since many of those numbers are negative.  

# 2.19
Attached to e-mail.



